---
title: "Assignment 2 - API Data Retrieval"
author: "Peter Menzies"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dataRetrieval)
library(tidyverse)
library(metajam)
library(here)
```

### Pull water data using USGS API and `dataRetrieval` package

```{r}
## Defining objects related to Ventura and Santa Paula Creek discharge based on USGS data-naming system
## Multiple sites can be requested by using a character vector for the sitNumbers argument
siteNumber <- c("11118500", "11113500")
parameterCd <- "00060"  # Discharge
startDate <- "2019-10-01"  
endDate <- "2021-10-07" 
```

```{r}
## Requesting data from USGS API using dataRetrieval::readNWISdv() and storing resultant dataframe
discharge <- readNWISdv(siteNumber, 
                    parameterCd, startDate, endDate) %>% 
  rename("discharge" = "X_00060_00003") %>% 
  mutate(site = case_when(
    site_no == "11118500" ~ "Ventura",
    site_no == "11113500" ~ "Santa_Paula"
  ))

discharge_ventura <- discharge %>% 
  filter(site == "Ventura") %>%
  filter(Date <= "2020-10-05")
```

```{r}
## Retrieving hourly discharge using readNWISuv() to find peak times after Oct 4 rain event
startDate2 <- "2021-10-04" 
endDate2 <- "2021-10-05"

discharge_oct4 <- readNWISuv(siteNumber, parameterCd, startDate2, endDate2) %>% 
  rename("discharge" = "X_00060_00000") %>% 
  mutate(site = case_when(
    site_no == "11118500" ~ "Ventura",
    site_no == "11113500" ~ "Santa_Paula"
  ))
```

### Plotting results 

```{r}
discharge_plot_oct4 <- ggplot(discharge_oct4, aes(x = dateTime, y = discharge)) +
  geom_line(aes(color = site))

discharge_plot_oct4

discharge_plot_ventura <- ggplot(discharge_ventura, aes(x = Date, y = discharge)) +
  geom_line()

discharge_plot_ventura

max_dis_vent <- discharge_oct4 %>% 
  filter(site_no == 11118500) %>% 
  filter(discharge == max(discharge))

max_dis_paula <- discharge_oct4 %>% 
  filter(site_no == 11113500) %>% 
  filter(discharge == max(discharge))
```
Based on the plot we created, and taking a look at the maximum discharge values around the Oct 4 rain event, it looks like discharge in Ventura peaked from around 7:30am to 5pm on Oct 5, whereas Santa Paula discharge peaked from around 10:30am to 11:15am that day.

### Downloading dataOne data using `metajam::download_d1_data()`

```{r, eval=FALSE}
## Storing URL to raw data and desired local file destination as objects to use as arguments in our download_d1_data() function
data_obj <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"
path <- "~/Downloads"
```

```{r, eval=FALSE}
## Requesting and downloading dataOne data using metajam
download_d1_data(data_obj, path)
```

```{r}
## Reading in contents of downloaded folder as list using metajam::read_d1_files()
household_language_list <- read_d1_files("/Users/petermenzies/MEDS/Fall_Quarter/EDS_213_Meta/doi_10.5063_F1CJ8BPH__household_language__csv 11-08-26-899.5063_F1CJ8BPH__household_language__csv")

household_language <- household_language_list$data

## Read in metadata
household_lang_metadata <- household_language_list$attribute_metadata

## Finding the average percent of English only speaking households in Alaska by year
household_lang_eng <- household_language %>% 
  filter(Year >= 2009 & Year <= 2015) %>% 
  group_by(Year) %>% 
  summarize(percent = (mean(speak_only_english, na.rm = T)) / (mean(total, na.rm = T)) * 100)

ggplot(household_lang_eng, aes(x = Year, y = percent)) +
  geom_line()

## Finding the average percent of English only speaking households by year and region (out of curiosity)
household_lang_eng_region <- household_language %>% 
  filter(Year >= 2009 & Year <= 2015) %>% 
  group_by(Year, SASAP.Region) %>% 
  summarize(percent = (mean(speak_only_english, na.rm = T)) / (mean(total, na.rm = T)) * 100)

ggplot(household_lang_eng_region, aes(x = Year, y = percent)) +
  geom_line(aes(color = SASAP.Region))
```

